---
title: "geospatial_correlation"
output: github_document
date: "2026-07-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Function summary 

Calcultes the correlation coefficient between two variables (numeric or logical) 
on an LSOA level and plots on a map. Plotting donewith library sf and ggplot2 and
colour pallette from  grDevices::colorRampPalette. 

Current testing done on synthetic health outcome data created by RVD. 

```{r, include=FALSE, echo=FALSE}

# retrieve LSOA boundaries for Cheshire and Merseyside
# to reduce coverage, remove names from the `within_names`

cm_lsoa_sf <- boundr::bounds(
  "lsoa",
  within_level = "lad",
  within_names = c(
    "Cheshire East",
    "Cheshire West and Chester",
    "Halton",
    "Knowsley",
    "Liverpool",
    "Sefton",
    "St. Helens",
    "Warrington",
    "Wirral"
  ),
  lookup_year = 2011, # can be either 2011 or 2021
  opts = boundr::boundr_options(resolution = "BFC")
)

liverpool_sf <- cm_lsoa_sf |>
  dplyr::filter(lad22nm %in% "Liverpool")

uprn_gs_sf <- "https://pldr.org/download/2k6r3/n12/UPRN_2_1_greenspace_distances_with_coords.csv" |>
  readr::read_csv() |>
  # convert to spatial object
  sf::st_as_sf(coords = c("longitude", "latitude"), crs = 4326) |>
  # add LSOA details
  sf::st_join(cm_lsoa_sf)



qof_asthma_tbl <- "https://pldr.org/download/e6nzv/ng1/QOF_4_03_Asthma_LSOA.csv" |>
  readr::read_csv() |>
  # filter out CM LSOAs
  dplyr::filter(lsoa11 %in% cm_lsoa_sf$lsoa11cd) |>
  # filter out last year
  dplyr::filter(year == 2024)



# set seed for reproducibility
set.seed(20250603)
qof_asthma_cohort_tbl <- qof_asthma_tbl |>
  purrr::pmap(function(lsoa11, den, num, ...) {
    # recalculate pop to be a multiple of 2
    pop <- ifelse(den %% 2 != 0, den + 1, den)
    # extract UPRNs for this LSOA
    uprn_tbl <- uprn_gs_sf |>
      sf::st_drop_geometry() |>
      dplyr::filter(lsoa11cd == lsoa11) |>
      dplyr::mutate(
        dist_gs_dec = dplyr::ntile(distance_local_greenspace, 10)
      ) |>
      dplyr::arrange(dist_gs_dec)
    
    # create cohort table
    cohort_tbl <- tibble::tibble(
      lsoa11cd = lsoa11,
      sex = rep(c(0, 1), each = pop / 2),
      has_asthma = FALSE,
      uprn = sample(uprn_tbl$UPRN, size = pop, replace = TRUE)
    )
    
    # randomly assigned the outcome to `num` patients
    idx <- sample(seq_len(pop), ceiling(num), replace = FALSE)
    cohort_tbl$has_asthma[idx] <- TRUE
    
    # randomly re-assigned UPRNs with higher distances to patients with outcome
    ## create vector of probabilities, so further distances are favoured
    p <- c(0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.1, 0.2, 0.2, 0.2)
    idx_uprn_tile <- sample(1:10, ceiling(num), replace = TRUE, prob = p)
    ## subset UPRNs with match tile as per `idx_uprn_tile`
    cohort_tbl$uprn[idx] <- purrr::map(idx_uprn_tile, function(x) {
      # subset based on tile decile  
      aux <- dplyr::filter(uprn_tbl, dist_gs_dec == x)
      # extract UPRN
      aux$UPRN[sample(seq_len(nrow(aux)), 1)]
    }) |>
      purrr::list_c()
    
    # return final cohort table
    return(cohort_tbl)
  }) |>
  purrr::list_c()


qof_asthma_cohort_dist_gs_sf <- qof_asthma_cohort_tbl |>
  dplyr::left_join(uprn_gs_sf, by = c("uprn" = "UPRN", "lsoa11cd")) |>
  dplyr::arrange(lsoa11cd, uprn) |>
  sf::st_as_sf()
```


# Upload data data and load back with DSLite  

```{r , include=FALSE, echo=FALSE}
    library(DSLite)
    library(devtools)
    library(dsBase)
    library(dsBaseClient)
    library(dsGeospatial)

    dat <- qof_asthma_cohort_dist_gs_sf |> 
      dplyr::select("lsoa11cd", "has_asthma","uprn","distance_local_greenspace") |> 
      sf::st_drop_geometry()

    devtools::load_all("/Users/tkariya/Documents/DataSHIELD/dsExample")

    ## ℹ Loading dsExample

    #devtools::load_all("/Users/tkariya/Documents/DataSHIELD/dsExampleClient")

    ## ℹ Loading dsExampleClient

    dslite.server1 <- newDSLiteServer(
      tables = list(
        uprn_asthma_green1 = dat
      )
    )

    dslite.server2 <- newDSLiteServer(
      tables = list(
        uprn_asthma_green2 = dat
      )
    )


    dslite.server1$config(defaultDSConfiguration(include=c("dsBase", "dsExample")))
    dslite.server1$aggregateMethod("groupcorDS", "groupcorDS")

    dslite.server2$config(defaultDSConfiguration(include=c("dsBase", "dsExample")))
    dslite.server2$aggregateMethod("groupcorDS", "groupcorDS")


    options(
      datashield.privacyControlLevel = "banana",
      nfilter.tab = 3,
      nfilter.subset = 3,
      nfilter.glm = 0.33,
      nfilter.string = 80,
      nfilter.stringShort = 20,
      nfilter.kNN = 3,
      nfilter.levels.density = 0.33,
      nfilter.levels.max = 40,
      nfilter.noise = 0.25,
      nfilter.privacy.old = 5
    )


    builder <- DSI::newDSLoginBuilder()

    builder$append(
      server = "server_1",
      url = "dslite.server1",
      table = "uprn_asthma_green1",
      driver = "DSLiteDriver"
    )

    builder$append(
      server = "server_2",
      url = "dslite.server2",
      table = "uprn_asthma_green2",
      driver = "DSLiteDriver"
    )


    logindata <- builder$build()

    conns <- DSI::datashield.login(logins = logindata, assign = TRUE)

```

Set nfilter.tab 

```{r, eval=FALSE}
nfilter.tab  <- 3
```


## Server-side function

```{r, eval=FALSE, echo=FALSE}
groupcorDS <- function (X, Y, INDEX){

  #############################################################
  #MODULE 1: CAPTURE THE nfilter SETTINGS
  thr <- dsBase::listDisclosureSettingsDS()
  nfilter.tab <- as.numeric(thr$nfilter.tab)
  nfilter.glm <- as.numeric(thr$nfilter.glm)
  #############################################################

  ## define corDS function

  FUN <- function(x){
    X <- x$X
    Y <- x$Y
    dataframe <- as.data.frame(cbind(X,Y))


    # names of the variables
    cls <- colnames(dataframe)

    # number of the input variables
    N.vars <- ncol(dataframe)

    ######################
    # DISCLOSURE CONTROLS
    ######################

    ##############################################################
    # FIRST TYPE OF DISCLOSURE TRAP - TEST FOR OVERSATURATION
    # TEST AGAINST nfilter.glm
    ##############################################################

    varcov.saturation.invalid <- 0

    if(N.vars > (nfilter.glm * nrow(dataframe))){

      varcov.saturation.invalid <- 1

      studysideMessage <- "ERROR: The ratio of the number of variables over the number of individual-level
                          records exceeds the allowed threshold, there is a possible risk of disclosure"
      stop(studysideMessage, call. = FALSE)

    }

    # CHECK X MATRIX VALIDITY
    # Check no dichotomous X vectors with between 1 and filter.threshold
    # observations at either level

    X.mat <- as.matrix(dataframe)

    dimX <- dim(X.mat)

    num.Xpar <- dimX[2]

    Xpar.invalid <- rep(0, num.Xpar)

    for(pj in 1:num.Xpar){
      unique.values.noNA <- unique((X.mat[,pj])[stats::complete.cases(X.mat[,pj])])
      if(length(unique.values.noNA)==2){
        tabvar <- table(X.mat[,pj])[table(X.mat[,pj])>=1] #tabvar COUNTS N IN ALL CATEGORIES WITH AT LEAST ONE OBSERVATION
        min.category <- min(tabvar)
        if(min.category < nfilter.tab){
          Xpar.invalid[pj] <- 1
        }
      }
    }

    # if any of the vectors in X matrix is invalid then the function returns an error

    if(is.element('1', Xpar.invalid)==TRUE & varcov.saturation.invalid==0){

      studysideMessage <- "ERROR: at least one variable is binary with one category less than the filter threshold for table cell size"
      stop(studysideMessage, call. = FALSE)

    }

    # if all vectors in X matrix are valid then the output matrices are calculated

    if(is.element('1', Xpar.invalid)==FALSE & varcov.saturation.invalid==0){

      # calculate the number of NAs in each variable separately
      column.NAs <- matrix(ncol=N.vars, nrow=1)
      colnames(column.NAs) <- cls
      for(i in 1:N.vars){
        column.NAs[1,i] <- length(dataframe[,i])-length(dataframe[stats::complete.cases(dataframe[,i]),i])
      }

      # remove any rows from the dataframe that include NAs
      casewise.dataframe <- dataframe[stats::complete.cases(dataframe),]

      # calculate the number of NAs casewise
      casewise.NAs.all <- dim(dataframe)[1]-dim(casewise.dataframe)[1]
      casewise.NAs <- matrix(casewise.NAs.all, ncol=N.vars, nrow=N.vars)
      rownames(casewise.NAs) <- cls
      colnames(casewise.NAs) <- cls

      # counts for NAs to be returned to the client:
      # This is a list with (a) a vector with the number of NAs in each variable (i.e. in each column)
      # separately and (b) the number of NAs casewise (i.e. the number of rows deleted from the input dataframe
      # which are the rows that at least one of their cells include a missing value)
      na.counts <- list(column.NAs, casewise.NAs)
      names(na.counts) <- list(paste0("Number of NAs in each column"), paste0("Number of NAs casewise"))

      # A matrix with elements the sum of products between each two variables
      sums.of.products <- matrix(ncol=N.vars, nrow=N.vars)
      rownames(sums.of.products) <- cls
      colnames(sums.of.products) <- cls
      for(m in 1:N.vars){
        for(p in 1:N.vars){
          sums.of.products[m,p] <- sum(as.numeric(as.character(casewise.dataframe[,m]))*as.numeric(as.character(casewise.dataframe[,p])))
        }
      }

      # A matrix with elements the sum of each variable
      sums <- matrix(ncol=N.vars, nrow=N.vars)
      rownames(sums) <- cls
      colnames(sums) <- cls
      for(m in 1:N.vars){
        for(p in 1:N.vars){
          sums[m,p] <- sum(as.numeric(as.character(casewise.dataframe[,m])))
        }
      }

      # A matrix with elements the sum of squares of each variable after removing missing values casewise
      sums.of.squares <- matrix(ncol=N.vars, nrow=N.vars)
      rownames(sums.of.squares) <- cls
      colnames(sums.of.squares) <- cls
      for(m in 1:N.vars){
        for(p in 1:N.vars){
          sums.of.squares[m,p] <- sum(as.numeric(as.character(casewise.dataframe[,m]))*as.numeric(as.character(casewise.dataframe[,m])))
        }
      }

      # A natrix with elements the number of complete cases casewise
      complete.counts <- matrix(dim(casewise.dataframe)[1], ncol=N.vars, nrow=N.vars)
      rownames(complete.counts) <- cls
      colnames(complete.counts) <- cls

    }

    return(list(sums.of.products=sums.of.products, sums=sums, complete.counts=complete.counts, na.counts=na.counts, sums.of.squares=sums.of.squares))
  }

  if(!identical(length(INDEX), length(X), length(Y))){
    stop("X, Y and INDEX should be of the same length")
  }


  analysis.matrix <- data.frame(X, Y, INDEX) # changed lines

  data.complete<-stats::complete.cases(analysis.matrix)

  Ntotal<-dim(analysis.matrix)[1]
  Nmissing<-sum(!data.complete)
  Nvalid<-sum(data.complete)

  simplify<-TRUE

  analysis.matrix.no.miss<-analysis.matrix[data.complete,]
  #nv<-dim(analysis.matrix)[2]

  # X<-as.vector(analysis.matrix.no.miss[,1])
  # Y<-as.vector(analysis.matrix.no.miss[,2])
  # INDEX <- analysis.matrix.no.miss[,3]

  dataframe <- as.data.frame(cbind(X,Y))

  if (!is.list(INDEX))
    INDEX <- list(INDEX)
  nI <- length(INDEX)
  if (!nI)
    stop("'INDEX' is of length zero")
  namelist <- vector("list", nI)
  names(namelist) <- names(INDEX)
  extent <- integer(nI)
  nx <- length(X)
  one <- 1L
  group <- rep.int(one, nx)
  ngroup <- one
  for (i in seq_along(INDEX)) {
    index <- as.factor(INDEX[[i]])
    if (length(index) != nx)
      stop("arguments must have same length")
    namelist[[i]] <- levels(index)
    extent[i] <- nlevels(index)
    group <- group + ngroup * (as.integer(index) - one)
    ngroup <- ngroup * nlevels(index)
  }



  #CALCULATE GROUP SIZES AND CHECK VALID

  ansmat.count<-table(group)

  # Set filter for cell sizes that are too small
  # the minimum number of observations that are allowed (the below function gets the value from opal)
  any.invalid.cell<-(sum(ansmat.count < nfilter.tab & ansmat.count > 0 ) >= 1)


  if(any.invalid.cell){
    table.valid<-FALSE
    cell.count.warning<-paste0("At least one group has between 1 and ", nfilter.tab-1, " observations. Please change groups")
    result<-list(table.valid,Nvalid,Nmissing,Ntotal,cell.count.warning)
    names(result)<-list("Table_valid","Nvalid","Nmissing","Ntotal","Warning")
    return(result)
  }


  if(!any.invalid.cell){
    table.valid<-TRUE
    cell.count.warning<-paste0("All tables valid")
    result <- lapply(split(dataframe, group), FUN) # check if to introduce lsoa names
    names(result) <- unlist(namelist)
    result[["Nfilter.tab"]] <- nfilter.tab
    return(result)
  }


}
```

## Client-side function
LSOA-level plotting can be done by combining or by splitting across the servers. 

```{r pressure, echo=FALSE, eval=FALSE}
#' @title Calculates the correlation of R objects in the server-side
#' @description This function calculates the correlation of two variables or the correlation
#' matrix for the variables of an input data frame.
#' @details In addition to computing correlations; this function produces a table outlining the
#' number of complete cases and a table outlining the number of missing values to allow the
#' user to decide the 'relevance' of the correlation based on the number of complete
#' cases included in the correlation calculations.
#'
#' If the argument \code{y} is not NULL, the dimensions of the object have to be
#' compatible with the argument \code{x}.
#'
#' The function calculates the pairwise correlations based on casewise complete cases which means that
#' it omits all the rows in the input data frame that include at least one cell with a missing value,
#' before the calculation of correlations.
#'
#' If \code{type} is set to \code{'split'} (default), the correlation of two variables or the
#' variance-correlation matrix of an input data frame and the number of complete cases and missing
#' values are returned for every single study. If type is set to \code{'combine'}, the pooled
#' correlation, the total number of complete cases and the total number of missing values aggregated
#' from all the involved studies, are returned.
#'
#' Server function called: \code{corDS}
#'
#' @param x a character string providing the name of the input vector, data frame or matrix.
#' @param y a character string providing the name of the input vector, data frame or matrix.
#' Default NULL.
#' @param type a character string that represents the type of analysis to carry out.
#' This must be set to \code{'split'} or \code{'combine'}.  Default \code{'split'}. For more information see details.
#' @param datasources a list of \code{\link[DSI]{DSConnection-class}} objects obtained after login.
#' If the \code{datasources} argument is not specified
#' the default set of connections will be used: see \code{\link[DSI]{datashield.connections_default}}.
#' @return \code{ds.cor} returns a list containing the number of missing values in each variable,
#' the number of missing variables casewise, the correlation matrix,
#' the number of used complete cases. The function applies two disclosure controls. The first disclosure
#' control checks that the number of variables is not bigger than a percentage of the individual-level records (the allowed
#' percentage is pre-specified by the 'nfilter.glm'). The second disclosure control checks that none of them is dichotomous
#' with a level having fewer counts than the pre-specified 'nfilter.tab' threshold.
#' @author DataSHIELD Development Team
#' @examples
#' \dontrun{
#'
#' ## Version 6, for version 5 see the Wiki
#'   # Connecting to the Opal servers
#'
#'   require('DSI')
#'   require('DSOpal')
#'   require('dsBaseClient')
#'
#'   builder <- DSI::newDSLoginBuilder()
#'   builder$append(server = "study1",
#'                  url = "http://192.168.56.100:8080/",
#'                  user = "administrator", password = "datashield_test&",
#'                  table = "CNSIM.CNSIM1", driver = "OpalDriver")
#'   builder$append(server = "study2",
#'                  url = "http://192.168.56.100:8080/",
#'                  user = "administrator", password = "datashield_test&",
#'                  table = "CNSIM.CNSIM2", driver = "OpalDriver")
#'   builder$append(server = "study3",
#'                  url = "http://192.168.56.100:8080/",
#'                  user = "administrator", password = "datashield_test&",
#'                  table = "CNSIM.CNSIM3", driver = "OpalDriver")
#'   logindata <- builder$build()
#'
#'   # Log onto the remote Opal training servers
#'   connections <- DSI::datashield.login(logins = logindata, assign = TRUE, symbol = "D")
#'
#'   # Example 1: Get the correlation matrix of two continuous variables
#'   ds.cor(x="D$LAB_TSC", y="D$LAB_TRIG", type="combine", datasources = connections)
#'
#'   # Example 2: Get the correlation matrix of the variables in a dataframe
#'   ds.dataFrame(x=c("D$LAB_TSC", "D$LAB_TRIG", "D$LAB_HDL", "D$PM_BMI_CONTINUOUS"),
#'                newobj="D.new", check.names=FALSE, datasources=connections)
#'   ds.cor("D.new", type="combine", datasources = connections)
#'
#'   # clear the Datashield R sessions and logout
#'   datashield.logout(connections)
#'
#' }
#' @export
#'
ds.groupcor <- function(x=NULL, y= NULL, index = NULL, type="split", datasources=NULL,
                        do.checks = TRUE, names_region = "Liverpool"){

  # look for DS connections
  if(is.null(datasources)){
    datasources <- datashield.connections_find()
  }

  # ensure datasources is a list of DSConnection-class
  if(!(is.list(datasources) && all(unlist(lapply(datasources, function(d) {methods::is(d,"DSConnection")}))))){
    stop("The 'datasources' were expected to be a list of DSConnection-class objects", call.=FALSE)
  }

  if(is.null(x)){
    stop("x=NULL. Please provide the name of a matrix or dataframe or the names of two numeric vectors!", call.=FALSE)
  }


  if(is.null(y)){
    stop("Please provide the name of the input vector!", call.=FALSE)
  }

  if(is.null(index)){
    stop("Please provide the name of the input vector!", call.=FALSE)
  }

  # check the type of the input objects
  if(do.checks){

    # check if the input objects are defined in all the studies
    isDefined(datasources, x)
    isDefined(datasources, y)

    # call the internal function that checks the input object is of the same class in all studies.
    typ1 <- checkClass(datasources, x)
    typ2 <- checkClass(datasources, y)
  }



  calltext <- paste0("groupcorDS(", x, ",", y, ",",index,")")

  output <- DSI::datashield.aggregate(datasources, as.symbol(calltext))

  # name of the studies to be used in the output
  stdnames <- names(datasources)

  ## Loading map
  if(type == 'split' && length(names_region) != length(stdnames)){
    names_region <- rep(names_region, times = length(stdnames))
  }

  shape_list <- lapply(names_region, function(region) {
    boundr::bounds(
      "lsoa",
      within_level = "lad",
      within_names = region,
      lookup_year = 2011,
      opts = boundr::boundr_options(resolution = "BFC")
    ) |>
      dplyr::select(lsoa11cd, geometry)
  })



  # call the server side function
  if (type=="split"){
    names(shape_list) <- stdnames
    covariance <- list()
    sqrt.diag <- list()
    correlation <- list()
    splitresults <- list()
    results <- list()
    for(i in 1:length(output)){
      for(j in 1: length(output[[i]])){
        covariance[[j]] <- matrix(0, ncol=dim(output[[i]][[j]][[1]])[2], nrow=dim(output[[i]][[j]][[1]])[1])
        correlation[[j]] <- matrix(0, ncol=dim(output[[i]][[j]][[1]])[2], nrow=dim(output[[i]][[j]][[1]])[1])
        colnames(correlation[[j]]) <- colnames(output[[i]][[j]][[1]])
        rownames(correlation[[j]]) <- colnames(output[[i]][[j]][[1]])
        for(m in 1:dim(output[[i]][[j]][[1]])[1]){
          for(n in 1:dim(output[[i]][[j]][[1]])[2]){
            covariance[[j]][m,n] <- (1/(output[[i]][[j]][[3]][m,n]-1))*(output[[i]][[j]][[1]][m,n])-(1/(output[[i]][[j]][[3]][m,n]*(output[[i]][[j]][[3]][m,n]-1)))*output[[i]][[j]][[2]][m,n]*output[[i]][[j]][[2]][n,m]
          }
        }
        correlation[[j]] <- stats::cov2cor(covariance[[j]])
        splitresults[[j]] <- list(output[[i]][[j]][[4]][[1]], output[[i]][[j]][[4]][[2]], correlation[[j]], output[[i]][[j]][[3]])
        n1 <- "Number of missing values in each variable"
        n2 <- "Number of missing values casewise"
        n3 <- "Correlation Matrix"
        n4 <- "Number of complete cases used"
        names(splitresults[[j]]) <- c(n1, n2, n3, n4)
      }
      names(splitresults) <- names(output[[i]])
      results[[i]] <- splitresults
    }
    names(results) <- stdnames

    results_data <- do.call(
      rbind,
      lapply(names(results), function(server_name) {

        server_results <- results[[server_name]]

        correlation_data <- data.frame(
          lsoa11cd = names(server_results),

          complete.counts = vapply(
            server_results,
            function(x) {
              x[["Number of complete cases used"]][1, 2]
            },
            numeric(1)
          ),

          correlation.coef = vapply(
            server_results,
            function(x) {
              x[["Correlation Matrix"]][1, 2]
            },
            numeric(1)
          ),

          server = server_name,
          row.names = NULL
        )

        dplyr::left_join(
          shape_list[[server_name]],
          correlation_data,
          by = "lsoa11cd"
        )
      })
    )

    results_data <- sf::st_as_sf(results_data)
    rownames(results_data) <- NULL

  }
  else if (type=="combine"){
      common_lsoas <- sort(Reduce(intersect, lapply(output, names)))

      if(is.null(common_lsoas) | length(common_lsoas) < output[[1]][["Nfilter.tab"]]){
        stop("Common lsoa in servers is less than the safe limit")
      }
      # lsoa in all servers should match or warning that common lsoa is only analysed
      # in that case check if the matrix size is not disclosive
      combined.results <- list()
      for (lsoa in common_lsoas) {

        reference <- output[[1]][[lsoa]]

         combined.sums.of.products <- matrix(
            0,
            nrow = nrow(reference[[1]]),
            ncol = ncol(reference[[1]]),
            dimnames = dimnames(reference[[1]])
          )

          combined.sums <- matrix(
            0,
            nrow = nrow(reference[[2]]),
            ncol = ncol(reference[[2]]),
            dimnames = dimnames(reference[[2]])
          )

          combined.complete.cases <- matrix(
            0,
            nrow = nrow(reference[[3]]),
            ncol = ncol(reference[[3]]),
            dimnames = dimnames(reference[[3]])
          )

          combined.missing.cases.vector <- matrix(
            0,
            nrow = nrow(reference[[4]][[1]]),
            ncol = ncol(reference[[4]][[1]]),
            dimnames = dimnames(reference[[4]][[1]])
          )

          combined.missing.cases.matrix <- matrix(
            0,
            nrow = nrow(reference[[4]][[2]]),
            ncol = ncol(reference[[4]][[2]]),
            dimnames = dimnames(reference[[4]][[2]])
          )

          combined.sums.of.squares <- matrix(
            0,
            nrow = nrow(reference[[5]]),
            ncol = ncol(reference[[5]]),
            dimnames = dimnames(reference[[5]])
          )

          # Combine sufficient statistics across relevant servers
          for (i in stdnames) {

            lsoa_output <- output[[i]][[lsoa]]

            combined.sums.of.products <-
              combined.sums.of.products + lsoa_output[[1]]

            combined.sums <-
              combined.sums + lsoa_output[[2]]

            combined.complete.cases <-
              combined.complete.cases + lsoa_output[[3]]

            combined.missing.cases.vector <-
              combined.missing.cases.vector + lsoa_output[[4]][[1]]

            combined.missing.cases.matrix <-
              combined.missing.cases.matrix + lsoa_output[[4]][[2]]

            combined.sums.of.squares <-
              combined.sums.of.squares + lsoa_output[[5]]
          }

          # Calculate covariance matrix for this LSOA
          combined.covariance <- matrix(
            NA_real_,
            nrow = nrow(combined.sums.of.products),
            ncol = ncol(combined.sums.of.products),
            dimnames = dimnames(combined.sums.of.products)
          )

          for (m in seq_len(nrow(combined.sums.of.products))) {
            for (n in seq_len(ncol(combined.sums.of.products))) {

              n_complete <- combined.complete.cases[m, n]

              if (!is.na(n_complete) && n_complete > 1) {

                combined.covariance[m, n] <-
                  combined.sums.of.products[m, n] / (n_complete - 1) -
                  (
                    combined.sums[m, n] *
                      combined.sums[n, m]
                  ) / (
                    n_complete * (n_complete - 1)
                  )
              }
            }
          }

          # Convert covariance to correlation
          combined.correlation <- tryCatch(
            stats::cov2cor(combined.covariance),
            error = function(e) {
              matrix(
                NA_real_,
                nrow = nrow(combined.covariance),
                ncol = ncol(combined.covariance),
                dimnames = dimnames(combined.covariance)
              )
            }
          )

          combined.results[[lsoa]] <- list(
            "Number of missing values in each variable" =
              combined.missing.cases.vector,

            "Number of missing values casewise" =
              combined.missing.cases.matrix,

            "Number of complete cases used" =
              combined.complete.cases,

            "Covariance Matrix" =
              combined.covariance,

            "Correlation Matrix" =
              combined.correlation
          )
        }

        results <- combined.results

        correlation_data <- data.frame(
          lsoa11cd = names(results),
          complete.counts = sapply(
            results,
            function(x) x$`Number of complete cases used`[1, 2]
          ),
          correlation.coef = sapply(
            results,
            function(x) x$`Correlation Matrix`[1, 2]
          ),
          server = if (type == "split") names(results) else rep("combine", length(results)),
          row.names = NULL
        )

        results_data <- dplyr::left_join(
            shape_list[[1]],
            correlation_data,
            by = "lsoa11cd"
          ) %>% dplyr::mutate(server = "combine")
      }

    else{
      stop('Function argument "type" has to be either "combine" or "split"', call.=FALSE)
    }

# Plotting
##### Load shapefile
 corplot <- ggplot2::ggplot(results_data) +
      ggplot2::geom_sf(ggplot2::aes(fill = correlation.coef)) +
      ggplot2::scale_fill_gradientn(
        colours = grDevices::colorRampPalette(c(
          "#440154",
          "#414487",
          "#2A788E",
          "#22A884",
          "#7AD151",
          "#FDE725"
        ))(100),
        na.value = "grey90"
      ) +
      ggplot2::labs(fill = "Correlation Coefficient") +
      ggplot2::facet_wrap(~server) +
      ggplot2::theme_minimal()


return(corplot)
}

```

## Output 

```{r}
ds.groupcor("D$has_asthma", "D$distance_local_greenspace", "D$lsoa11cd", datasources = conns, type = "combine")
ds.groupcor("D$has_asthma", "D$distance_local_greenspace", "D$lsoa11cd", datasources = conns, type = "split")
```



